### 2025-02-27 Early_descriptives.tex
Updates from Overleaf

\documentclass[12pt]{article}

\usepackage{tgtermes}
\usepackage{epsf}
\usepackage{epstopdf}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage[colorlinks=true,linkcolor=blue,citecolor=blue]{hyperref}
\usepackage{dcolumn}
\usepackage{amsmath, amsthm, amssymb}
\usepackage{mwe}
\usepackage{url}
%\usepackage{harvard}
\usepackage{fancyhdr}
\usepackage{longtable}
\usepackage{authblk}
\usepackage{setspace}
%\usepackage[nomarkers]{endfloat}
\usepackage{float}
\usepackage{bbm}
%\usepackage{titling}
\usepackage{subcaption}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{import}
\usepackage[backend=biber,style=authoryear,
sorting=ynt,citestyle=authoryear]{biblatex}
\addbibresource{papercitations.bib}
%\usepackage[nomarkers,nofiglist,notablist]{endfloat}
\usepackage{subcaption}
\usepackage{caption}

\onehalfspacing
\textwidth 6.5in \oddsidemargin 0in \evensidemargin -0.6in
\textheight 8.5in \topmargin -0.2in

\newcolumntype{L}[1]{>{\raggedright\let\newline\\
		\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{C}[1]{>{\centering\let\newline\\
		\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{R}[1]{>{\raggedleft\let\newline\\
		\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{P}[1]{>{\raggedright\tabularxbackslash}p{#1}}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}

\captionsetup{justification=centering,singlelinecheck=false}

\begin{document}

	\linespread{1.2}\title{\vspace{-0.5in} Common Ownership on Boards\\ \large Evidence from Nonprofit Hospitals}

	\date{\today}

	\author{\vspace{10mm}Hanna Glenn\footnote{Department of Economics, Emory University, 1602 Fishburne Drive, Atlanta, GA 30322, hanna.glenn@emory.edu.} }

	\maketitle

	\singlespacing

    \vspace{3mm}

    \begin{abstract}
		{\small

		} 
	\end{abstract}

	\onehalfspacing

	\newpage

    \section{Early Descriptives: Connected in System}

    \begin{figure}[ht!]
        \centering
        \includegraphics[width=\textwidth]{Objects/connected_systems_percent.pdf}
    \end{figure}

    \newpage

    \begin{figure}[ht!]
    \centering
        \includegraphics[width=.9\textwidth]{Objects/common_boards_insystem_maps.pdf}
        \label{fig:outcomes_graph}
    \end{figure}

    \import{Objects}{means_insystem_table.tex}

    \import{Objects}{connected_syspair_diff_table.tex}

    \newpage

    \section{Early Descriptives: Independent Connected Hospitals}

    \begin{figure}[ht!]
        \centering
        \includegraphics[width=\textwidth]{Objects/connected_independent_percent.pdf}
    \end{figure}

    \begin{figure}[ht!]
    \centering
        \includegraphics[width=\textwidth]{Objects/common_boards_independent_maps.pdf}
        \label{fig:outcomes_graph}
    \end{figure}

    \import{Objects}{means_independent_table.tex}

    \import{Objects}{connected_indpair_diff_table.tex}

\end{document}